A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2020-12-17 cusp_j/paper/paper.tex
conclusion almost finished

\begin{abstract}
TODO: investigate the non variationality on the He atom for instance and Ne8+

@@

\subsection{A simple physical picture: the transcorrelated Hamiltonian for the helium atom}
Let us write the Hamiltonian of the helium atom using the $r_i = \big|{\bf r}_i\big|$ and $ r_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinates  
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg).
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the transcorrelated Hamiltonian $\tilde{H}[u]$ obtained by a similarity transformation of the usual Hamiltonian by a Jastrow factor $u(r_{12})$ depending only on $r_{12}$: 
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[u]&= e^{-u(r_{12})} H e^{u(r_{12})}.
 \end{aligned}
\end{equation}
The only additional terms arising in $\tilde{H}[u]$ with respect to $H$ are those coming from the action of the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[u] =  -e^{-u(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{u(r_{12})}.  
\end{equation}
By defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[u] = -2 \deriv{u(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[u] = -\frac{2}{r_{12}} \deriv{u(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[u] = -\deriv{u(r_{12})}{r_{12}}{2} - \bigg( \deriv{u(r_{12})}{r_{12}}{} \bigg)^2, 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_general}
 \tilde{H}[u] = h_c + \frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u] + \tilde{t}[u].
\end{equation}
Therefore, one can see from Eq. \eqref{eq:ht_general} that the two electrons experience a modified scalar potential given by $\frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u]$, and an additional differential operator $\tilde{t}[u]$, 
the latter making $\tilde{H}[u]$ non-hermitian. It is important to notice that the fact that  $\tilde{H}[u]$ is non-hermitian implies the loose of variational principle.

@@

\subsection{The general analytical form of  $\tilde{H}[\mu]$ and some of its properties}
\label{sec:ht_general}
\subsubsection{General form of $\tilde{H}[\mu]$ }
Having established the analytical expression of $u(r_{12};\mu)$, one can derive the general expression of the associated transcorrelated Hamiltonian.  
According to Eq. (2) of Ref. \onlinecite{CohLuoGutDobTewAla-JCP-19}, the similarity transformed Hamiltonian can be written as 
\begin{equation}
 \label{ht_def_g}
 e^{-\hat{\tau}} \hat{H} e^{\hat{\tau}} = H + \big[ H,\hat{\tau} \big] + \frac{1}{2}\bigg[ \big[H,\hat{\tau}\big],\hat{\tau}\bigg]
\end{equation}
where $\hat{\tau} = \sum_{i<j}u(\br{}_i,\br{}_j)$ and $\hat{H} = \sum_i -\frac{1}{2} \nabla^2_i + v(\br{}_i) + \sum_{i<j} \frac{1}{r_{ij}}$. 
This leads to the following similarity transformed Hamiltonian 
\begin{equation}
 \begin{aligned}
 \label{ht_def_gu}
 \tilde{H} & = H - \sum_i \bigg( \frac{1}{2} \nabla_i^2 \hat{\tau} + \big(\nabla_i \hat{\tau} \big) + \frac{1}{2} \big(\nabla_i \hat{\tau} \big)^2  \bigg) \\
           & = H - \sum_{i<j} \hat{K}(\bri{i},\bri{u}) - \sum_{i<j<k} \hat{L}(\bri{i},\bri{u},\bri{k}),
 \end{aligned}
\end{equation}
where the effective two- and three-body operators $\hat{K}(\bri{i},\bri{u})$ and $\hat{L}(\bri{i},\bri{u},\bri{k})$ are defined as
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{K}(\bri{1},\bri{2}) ) = \frac{1}{2} \bigg( &\nabla_1^2 u(\bri{1},\bri{2}) + \nabla_2^2u(\bri{1},\bri{2}) \\
                                               + &\big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 \bigg) \\
                                               + &\nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_2 + \nabla_2 u(\bri{1},\bri{2}) \cdot \nabla_1,
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}(\bri{1},\bri{2},\bri{3}) ) = & \nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_1 u(\bri{1},\bri{3}) + \nabla_2 u(\bri{2},\bri{1}) \cdot \nabla_2 u(\bri{2},\bri{3})  \\
                                     + & \nabla_3 u(\bri{3},\bri{1}) \cdot \nabla_3 u(\bri{3},\bri{2}).
 \end{aligned}
\end{equation}
Therefore, to obtain the analytical expression of the $\tilde{H}[\mu]$ for a general $N$-electron system, 
one needs to compute the actual form of the operators $\hat{K}(\bri{1},\bri{2})$ and $\hat{L}(\bri{1},\bri{2},\bri{3})$ when $u(\bri{1},\bri{2}) = u(r_{12},\mu)$.

\subsection{Analytical form for the operators $\hat{K}(\bri{1},\bri{2}) $ and $\hat{L}(\bri{1},\bri{2},\bri{3})$ }
To compute the operators $\hat{K}(\bri{1},\bri{2})$ and $\hat{L}(\bri{1},\bri{2},\bri{3}) $ one needs to compute first the gradient of $u(r_{12},\mu)$ with respect to $\bri{1}$ which reads 
\begin{equation}
 \label{eq:nabla_u}
 \begin{aligned}
 \nabla_1 u(\br{}_1,\br{}_2) = \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \big( \br{}_1 - \br{}_2 \big).
 \end{aligned}
\end{equation}
Then, the term $\big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2$ is simply 
\begin{equation}
 \begin{aligned}
 \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_2 u(\bri{1},\bri{2}) \big) ^2 = \frac{\bigg(1 - \text{erf}(\mu r_{12}) \bigg)^2}{2}.
 \end{aligned}
\end{equation}
Then, the non hermitian operators in $\hat{K}(\bri{1},\bri{2})$ contains 
\begin{equation}
 \begin{aligned}
 \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1  = &\frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \\ 
                                          &\bigg( (x_1 - x_2) \deriv{}{x_1}{} + (y_1 - y_2) \deriv{}{y_1}{} + (z_1 - z_2) \deriv{}{z_1}{}\bigg),
 \end{aligned}
\end{equation}
and as $\nabla_1 u(\br{}_1,\br{}_2) = - \nabla_2 u(\br{}_1,\br{}_2)$,   
the total non hermitian operator in $\hat{K}(\bri{1},\bri{2})$ can be written as 
\begin{equation}
 \begin{aligned}
 \label{def_non_hermit}
& \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1 + \nabla_2 u(\br{}_1,\br{}_2) \cdot \nabla_2 = \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \\
& \bigg( (x_1 - x_2) \big( \deriv{}{x_1}{} - \deriv{}{x_2}{} \big) +
         (y_1 - y_2) \big( \deriv{}{y_1}{} - \deriv{}{y_2}{} \big)  \\
&  +      (z_1 - z_2) \big( \deriv{}{z_1}{} - \deriv{}{z_2}{} \big)\bigg).
 \end{aligned}
\end{equation}
One can notice that the form of Eq. \eqref{def_non_hermit} provides an explicit form to compute analytical integrals with Gaussian types functions. 
Nevertheless, one can notice that as 
\begin{equation}
 \deriv{}{r_{12}^x}{} = \frac{1}{2} \bigg( \deriv{}{x_1}{} - \deriv{}{x_2}{} \bigg),
\end{equation}
one can write 
\begin{equation}
 \begin{aligned}
 \label{eq:nabla_i_nabla0}
& \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1 + \nabla_2 u(\br{}_1,\br{}_2) \cdot \nabla_2 = \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} \big( \br{}_1 - \br{}_2 \big) \cdot \nabla_{\br{}_{12}}.
 \end{aligned}
\end{equation}
Then, introducing the spherical coordinate system for $\br{}_1 - \br{}_2$ in terms of ${\bf e}_u= \frac{\br{}_1 - \br{}_2}{r_{12}}$ and the angle $\theta$ and $\phi$, one can write 
\begin{equation}
 \nabla_{\br{}_{12}} = \deriv{}{r_{12}}{} {\bf e}_u + \frac{1}{r_{12}} \deriv{}{\theta}{} {\bf e}_{\theta} + \frac{1}{r_{12} \sin(\theta)} \deriv{}{\phi}{} {\bf e}_\phi,
\end{equation}
and as $\br{}_1 - \br{}_2 = r_{12} {\bf e}_u$ one obtains
\begin{equation}
 \big( \br{}_1 - \br{}_2 \big) \cdot \nabla_{\br{}_{12}} = r_{12} \deriv{}{r_{12}}{},
\end{equation}
and therefore 
\begin{equation}
 \begin{aligned}
 \label{eq:nabla_i_nabla1}
 \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1 + \nabla_2 u(\br{}_1,\br{}_2) \cdot \nabla_2 = \bigg( 1 - \text{erf}(\mu r_{12})\bigg) \deriv{}{r_{12}}{},
 \end{aligned}
\end{equation}
which is a more compact expression than Eq.\eqref{def_non_hermit}

The computation of the Laplacian is a little more tedious, but after some derivations it comes  
\begin{equation}
 \begin{aligned}
 \label{eq:d2_x1_2}
 &\nabla_1^2 u(\bri{1},\bri{2}) + \nabla_2^2u(\bri{1},\bri{2})\\ 
 & = 2 \times \bigg( \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} - \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2}  \bigg).
 \end{aligned}
\end{equation}

@@

\section{Computation of integrals}

>>>>

\begin{abstract}
The present work introduces a new form of explicitly correlated factor in the context of the transcorrelated methods. 
The new correlation factor is obtained from the $r_{12} \approx 0$ mathematical analysis of the transcorrelated Hamiltonian, and its analytical form is obtained such that the scalar part of the effective two-electron potential reproduces at leading order in $1/r_{12}$ the long-range interaction of the range-separated density functional theory. 
The resulting correlation factor is then tuned by a unique parameter $\mu$ which controls both the depth of the coulomb hole and its typical range in $r_{12}$. 
The transcorrelated Hamiltonian obtained with such a new correlation factor as a straightforward analytical expression and its physical contents depends on the same parameter $\mu$: one can continuously change from a non divergent repulsive Hamiltonian potential at large $\mu$ to a purely attractive one at small $\mu$. We investigate the convergence of the eigenvalues and right-eigenvectors of such new transcorrelated Hamiltonian as a function of the basis set and as a function of $\mu$ on a series of two-electron systems. 
We found that the convergence towards the complete basis set is much faster for a quite wide range values of $\mu$. We also propose a computationally cheap way to find a reasonable value of $\mu$ that automatically adapts to the system through its density.  
It should be pointed out that all two-electron integrals can be carried analytically within such approach and the three-electrons terms can be performed using a mixed analytical-numerical integration approach.

@@

\subsection{A simple physical picture: the transcorrelated Hamiltonian for the helium atom}
\label{sec:he_htilde}
Let us write the Hamiltonian of the helium atom using the $r_i = \big|{\bf r}_i\big|$ and $ r_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinates  
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg).
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the transcorrelated Hamiltonian $\tilde{H}[u]$ obtained by a similarity transformation of the usual Hamiltonian by a Jastrow factor $u(r_{12})$ depending only on $r_{12}$: 
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[u]&= e^{-u(r_{12})} H e^{u(r_{12})}.
 \end{aligned}
\end{equation}
The only additional terms arising in $\tilde{H}[u]$ with respect to $H$ are those coming from the action of the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[u] =  -e^{-u(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{u(r_{12})}.  
\end{equation}
By defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[u] = -2 \deriv{u(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[u] = -\frac{2}{r_{12}} \deriv{u(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[u] = -\deriv{u(r_{12})}{r_{12}}{2} - \bigg( \deriv{u(r_{12})}{r_{12}}{} \bigg)^2, 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_general}
 \tilde{H}[u] = h_c + \frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u] + \tilde{t}[u].
\end{equation}
Therefore, one can see from Eq. \eqref{eq:ht_general} that the two electrons experience a modified scalar potential given by $\frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u]$, and an additional differential operator $\tilde{t}[u]$, 
the latter making $\tilde{H}[u]$ non-hermitian. It is important to notice that the fact that  $\tilde{H}[u]$ is non-hermitian implies the loose of variational principle.

@@

\subsection{The general analytical form of  $\tilde{H}[\mu]$ and some of its properties}
\label{sec:ht_general}
\subsubsection{General form of $\tilde{H}[\mu]$ }
Having established the analytical expression of $u(r_{12};\mu)$, one can derive the general expression of the associated transcorrelated Hamiltonian.  
According to Eq. (2) of Ref. \onlinecite{CohLuoGutDobTewAla-JCP-19}, the similarity transformed Hamiltonian can be written as 
\begin{equation}
 \label{ht_def_g}
 e^{-\hat{\tau}} \hat{H} e^{\hat{\tau}} = H + \big[ H,\hat{\tau} \big] + \frac{1}{2}\bigg[ \big[H,\hat{\tau}\big],\hat{\tau}\bigg]
\end{equation}
where $\hat{\tau} = \sum_{i<j}u(\br{}_i,\br{}_j)$ and $\hat{H} = \sum_i -\frac{1}{2} \nabla^2_i + v(\br{}_i) + \sum_{i<j} \frac{1}{r_{ij}}$. 
This leads to the following similarity transformed Hamiltonian 
\begin{equation}
 \begin{aligned}
 \label{ht_def_gu}
 \tilde{H} & = H - \sum_i \bigg( \frac{1}{2} \nabla_i^2 \hat{\tau} + \big(\nabla_i \hat{\tau} \big) + \frac{1}{2} \big(\nabla_i \hat{\tau} \big)^2  \bigg) \\
           & = H - \sum_{i<j} \hat{K}(\bri{i},\bri{j}) - \sum_{i<j<k} \hat{L}(\bri{i},\bri{j},\bri{k}),
 \end{aligned}
\end{equation}
where the effective two- and three-body operators $\hat{K}(\bri{i},\bri{j})$ and $\hat{L}(\bri{i},\bri{j},\bri{k})$ are defined as
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{K}(\bri{1},\bri{2}) ) = \frac{1}{2} \bigg( &\nabla_1^2 u(\bri{1},\bri{2}) + \nabla_2^2u(\bri{1},\bri{2}) \\
                                               + &\big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 \bigg) \\
                                               + &\nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_2 + \nabla_2 u(\bri{1},\bri{2}) \cdot \nabla_1,
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}(\bri{1},\bri{2},\bri{3}) ) = & \nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_1 u(\bri{1},\bri{3}) + \nabla_2 u(\bri{2},\bri{1}) \cdot \nabla_2 u(\bri{2},\bri{3})  \\
                                     + & \nabla_3 u(\bri{3},\bri{1}) \cdot \nabla_3 u(\bri{3},\bri{2}).
 \end{aligned}
\end{equation}
Therefore, to obtain the analytical expression of the $\tilde{H}[\mu]$ for a general $N$-electron system, 
one needs to compute the actual form of the operators $\hat{K}(\bri{1},\bri{2})$ and $\hat{L}(\bri{1},\bri{2},\bri{3})$ when $u(\bri{1},\bri{2}) = u(r_{12},\mu)$.

\subsection{Analytical form for the operators $\hat{K}(\bri{1},\bri{2}) $ and $\hat{L}(\bri{1},\bri{2},\bri{3})$ }
To compute the operators $\hat{K}(\bri{1},\bri{2})$ and $\hat{L}(\bri{1},\bri{2},\bri{3}) $ one needs to compute first the gradient of $u(r_{12},\mu)$ with respect to $\bri{1}$ which reads 
\begin{equation}
 \label{eq:nabla_u}
 \begin{aligned}
 \nabla_1 u(\br{}_1,\br{}_2) = \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \big( \br{}_1 - \br{}_2 \big).
 \end{aligned}
\end{equation}
Then, the term $\big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2$ is simply 
\begin{equation}
 \begin{aligned}
 \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_2 u(\bri{1},\bri{2}) \big) ^2 = \frac{\bigg(1 - \text{erf}(\mu r_{12}) \bigg)^2}{2}.
 \end{aligned}
\end{equation}
The non hermitian operators in $\hat{K}(\bri{1},\bri{2})$ contains 
\begin{equation}
 \begin{aligned}
 \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1  = &\frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \\ 
                                          &\bigg( (x_1 - x_2) \deriv{}{x_1}{} + (y_1 - y_2) \deriv{}{y_1}{} + (z_1 - z_2) \deriv{}{z_1}{}\bigg),
 \end{aligned}
\end{equation}
and as $\nabla_1 u(\br{}_1,\br{}_2) = - \nabla_2 u(\br{}_1,\br{}_2)$,   
the total non hermitian operator in $\hat{K}(\bri{1},\bri{2})$ can be written as 
\begin{equation}
 \begin{aligned}
 \label{def_non_hermit}
& \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1 + \nabla_2 u(\br{}_1,\br{}_2) \cdot \nabla_2 = \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \\
& \bigg( (x_1 - x_2) \big( \deriv{}{x_1}{} - \deriv{}{x_2}{} \big) +
         (y_1 - y_2) \big( \deriv{}{y_1}{} - \deriv{}{y_2}{} \big)  \\
&  +      (z_1 - z_2) \big( \deriv{}{z_1}{} - \deriv{}{z_2}{} \big)\bigg).
 \end{aligned}
\end{equation}
One can notice that the form of Eq. \eqref{def_non_hermit} provides an explicit form to compute analytical integrals with Gaussian types functions. 
Nevertheless, one can notice that as 
\begin{equation}
 \deriv{}{r_{12}^x}{} = \frac{1}{2} \bigg( \deriv{}{x_1}{} - \deriv{}{x_2}{} \bigg),
\end{equation}
one can write 
\begin{equation}
 \begin{aligned}
 \label{eq:nabla_i_nabla0}
& \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1 + \nabla_2 u(\br{}_1,\br{}_2) \cdot \nabla_2 = \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} \big( \br{}_1 - \br{}_2 \big) \cdot \nabla_{\br{}_{12}}.
 \end{aligned}
\end{equation}
Then, introducing the spherical coordinate system for $\br{}_1 - \br{}_2$ in terms of ${\bf e}_u= \frac{\br{}_1 - \br{}_2}{r_{12}}$ and the angle $\theta$ and $\phi$, one can write 
\begin{equation}
 \nabla_{\br{}_{12}} = \deriv{}{r_{12}}{} {\bf e}_u + \frac{1}{r_{12}} \deriv{}{\theta}{} {\bf e}_{\theta} + \frac{1}{r_{12} \sin(\theta)} \deriv{}{\phi}{} {\bf e}_\phi,
\end{equation}
and as $\br{}_1 - \br{}_2 = r_{12} {\bf e}_u$ one obtains
\begin{equation}
 \big( \br{}_1 - \br{}_2 \big) \cdot \nabla_{\br{}_{12}} = r_{12} \deriv{}{r_{12}}{},
\end{equation}
and therefore 
\begin{equation}
 \begin{aligned}
 \label{eq:nabla_i_nabla1}
 \nabla_1 u(\br{}_1,\br{}_2) \cdot \nabla_1 + \nabla_2 u(\br{}_1,\br{}_2) \cdot \nabla_2 = \bigg( 1 - \text{erf}(\mu r_{12})\bigg) \deriv{}{r_{12}}{},
 \end{aligned}
\end{equation}
which is a more compact expression than Eq.\eqref{def_non_hermit}

The computation of the Laplacian is a little more tedious, but after some derivations it comes  
\begin{equation}
 \begin{aligned}
 \label{eq:d2_x1_2}
 &\nabla_1^2 u(\bri{1},\bri{2}) + \nabla_2^2u(\bri{1},\bri{2})
 & = 2 \times \bigg( \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} - \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2}  \bigg).
 \end{aligned}
\end{equation}

@@

\section{Appendix}
\subsection{$r_{12} \rightarrow 0$ limits for different forms of Hamiltonians and related cusp conditions}
\label{sec:cusp}
Having established the analytical form of the transcorrelated Hamiltonian $\tilde{H}[u]$ for a general Jastrow factor $u(r_{12})$ in the case of the helium atom in Sec. \ref{sec:he_htilde}, one can then study the behaviour of $\tilde{H}[u]$ when $r_{12}\rightarrow 0$ and compare it to two other types of Hamiltonian: the usual physical Hamiltonian and that entering the RS-DFT framework. 
The analysis of the $r_{12}\rightarrow 0$ behaviour leads to the different condition that the eigenvectors of such operators must fulfill. 
It should be noticed that, although it was established in the case of the helium atom, the leading order of the $r_{12} \rightarrow 0$ limit that will be carried out in that section are valid for a general $N$ electron system . 
For the sake of simplicity of the notations we focus on the ground state of these operators but the arguments are valid for any bounded eigenvectors.

Another form of effective Hamiltonians leading to cuspless eigenfunctions are those obtained from RS-DFT 
where the Coulomb interaction $1/r_{12}$ is split into a non-divergent long-range interaction $\text{erf}(\mu r_{12})/r_{12}$ and a complementary short-range interaction $\text{erfc}(\mu r_{12})/r_{12}$, where $\text{erf}(x)$ and $\text{erfc}(x)$ are the error and complementary error functions, respectively. 
The parameter which controls such a splitting is the so-called range separation parameter $\mu$:  RS-DFT reduces the usual WFT when $\mu \rightarrow \infty$, and $\mu \rightarrow 0$ it gives back the usual Kohn-Sham theory.  
In practice, RS-DFT introduces a self-consistent Schroedinger-like equation which must be fulfilled by an effective wave function  $\Psi^\mu$.
Applied to a two electron system and looking in regions where $r_{12}\approx 0$, the leading terms of the self-consistent Schroedinger-like equation of RS-DFT reads 
\begin{equation}
 \label{eq:cusp_psi_mu_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{\text{erf}(\mu r_{12})}{r_{12}}\bigg) \Psi^\mu(\bd{r_1},\bd{r_2})  = c,\quad c<\infty,  
\end{equation}
and 
\begin{equation}
 \lim_{r_{12} \rightarrow 0} \frac{\text{erf}(\mu r_{12})}{r_{12}} = \frac{2 \mu}{\sqrt{\pi}} , 
\end{equation}
one obtains that 
\begin{equation}
 \label{eq:cusp_psi_mu_1}
 \deriv{\Psi^\mu(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = 0, 
\end{equation}
which means that the wave functions $\Psi^\mu$, just as $\phiex$ are cuspless, because they deal with an effective non-divergent interaction.

\subsection{Large $\mu$ limit of $\tilde{H}[\mu]$}
\label{sec:large_mu_lim}
As mentioned in Sec. \ref{sec:h_mu_lim}, one must recover the usual physical Hamiltonian in the large $\mu$ limit. 
Nevertheless, one can notice that 
\begin{equation}
 \label{eq:lim_mu_3}
 \lim_{\mu \rightarrow \infty} \tilde{\mathcal{W}}_{ee}(r_{12})  = \frac{1}{r_{12}} + \delta(r_{12}) 
\end{equation}
with 
\begin{equation}
 \delta(x) = \lim_{\mu} \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu x \big)^2}, 
\end{equation}
being the Dirac distribution. Threfore, one could be tempted to write that 
\begin{equation}
 \begin{aligned}
 \label{eq:lim_mu_4}
 \lim_{\mu \rightarrow \infty} \tilde{H}[\mu]& = H + \delta(r_{12})  \\
                                             & \ne H,
 \end{aligned}
\end{equation}
which seems to contradict the intuitive limit of Eq.\eqref{eq:lim_mu_1}.

Nevertheless, as Eq.\eqref{eq:lim_mu_4} deals with operators, the equality must be considered in the sense of distributions.  
Therefore, considering a generic bounded square integrable wave function $\psi(\br{}_1,\hdots,\br{}_N)$ yielding to a integrable pair density $n_2(\br{}_1,\br{}_2)$, the equality between $\tilde{H}[\mu]$ and $H$ is, in the sense of distributions, defined by the following equality 
\begin{equation}
 \label{eq:lim_mu_5}
 \begin{aligned}
& \matelem{\psi}{H}{\psi} = \lim_{\mu \rightarrow \infty} \matelem{\psi}{\tilde{H}[\mu]}{\psi} \\
& \Leftrightarrow H = \tilde{H}[\mu],
 \end{aligned}
\end{equation}
and considering the limit for of $\tilde{H}[\mu]$ (see Eq.\eqref{eq:lim_mu_4}), it implies that 
\begin{equation}
 \label{eq:lim_mu_6}
 \int \text{d}\br{}_1 \text{d}\br{}_2 n_2(\br{}_1,\br{}_2) \delta(r_{12}) = 0.
\end{equation}
To intuitively show Eq.\eqref{eq:lim_mu_6}, one can perform a series of change of variable $(\br{}_1,\br{}_2)\rightarrow (\br{}_{12},\frac{1}{2}(\br{}_1 + \br{}_2))$ and then use the spherical coordinates for $\br{}_{12}$ 
\begin{equation}
 \int \text{d}\br{}_1 \text{d}\br{}_2 n_2(\br{}_1,\br{}_2) \delta(r_{12}) = \int \text{d}r{}_{12}  \tilde{n}_2(r_{12}) \delta(r_{12}) r_{12}^2 
\end{equation}
where $\tilde{n_2}(x)$ is the function $n_2(\br{}_1,\br{}_2)$ integrated over $(\br{}_1,\br{}_2)$ with the constraint that 
$|\br{}_1-\br{}_2| = x$. 
As the pair density $n_2(\br{}_1,\br{}_2)$ remains finite and is integrable, the function $\tilde{n}_2(x)$ 
cannot diverge faster than $\frac{1}{x}$ when $x\rightarrow 0$ and therefore one obtains that 
\begin{equation}
 \int \text{d}r{}_{12}  \tilde{n}_2(r_{12}) \delta(r_{12}) r_{12}^2 = 0,
\end{equation}
which implies Eq.\eqref{eq:lim_mu_1}.
